---
title: "Data Challenge 4: Monty Hall Simulation"
author: "Ziyan Hu"
date: "Due 5:59pm on Nov 6, 2023"
output:
  html_document:
---

Link to the github: <https://github.com/freyahu2000/DataChallenge4>

## Question 1
First write a function that does one play of the Monty Hall simulation. Have an argument `switch` which tells if the contestant uses the strategy of switching the door `(switch = TRUE)` or `(switch = FALSE)`.

- First use `sample()` to select the door that the contestant picks and the door that the prize is under
- Next use an if else statement to show what door is revealed if (1) the contestant guessed the correct door or (2) if the contestant did not guess the correct door
- Use an if statement to switch guesses of the door if `switch = TRUE`
- return a logical (TRUE/FALSE) indicating if the game was won or not

### Solution
```{r}
set.seed(1234)  
MontyHallSimulation <- function(switch){
  # initialize doors
  doors <- 1:3
  # contestant's choice
  contestant_pick <- sample(doors,1)
  # prize's location
  prize_door <- sample(doors,1)
  # if contestant chose the prize door
  if (contestant_pick == prize_door){
    # doors excluding contestant's pick
    remaining_doors <- doors[doors != contestant_pick]
    # door host reveals
    reveal_door <- sample(remaining_doors,1)
    # if contestant switches, they lose
    if(switch == TRUE){
      return(FALSE)
    }
  # if contestant didn't choose the prize door
  } else {
    # door host reveals
    reveal_door <- doors[doors != contestant_pick 
                         & doors != prize_door]
    # if contestant doesn't switch, they lose
    if(switch == FALSE){
      return(FALSE)
    }
  }
  # if none of the above, contestant win
  return(TRUE)
}

# one time simulation 
MontyHallSimulation(TRUE)
```

## Question 2
Next write a function that performs the simulation multiple times. Have an argument for the number of times to run the simulation `(n)` and the strategy `(switch)`. Check out the `replicate()` function which may be useful.

### Solution
```{r}
set.seed(1234)
MontyHallSimulationN <- function(n,switch){
  # perform monty hall simulation n times
  results <- replicate(n,MontyHallSimulation(switch))
  return(results)
}
```

```{r}
# simulate 1000 times with switch strategy
switchSimulation <- MontyHallSimulationN(1000, TRUE)
# winning probability with switch
switchWinProb <- sum(switchSimulation == TRUE)/1000 
switchWinProb
# simulate 1000 times without switch
noSwitchSimulation <- MontyHallSimulationN(1000, FALSE)
# winning probability without switch
noSwitchProb <- sum(noSwitchSimulation == TRUE)/1000 
noSwitchProb
```

Therefore, from the 1000 times simulations we can see that switching door strategy has greater probability to win the prize.
